### 2020-05-28 writeup/transfer.tex
sec 4

\begin{proposition}\label{prop_data_size}
	In the simplified model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
\item If $d^2 \le \frac{\sigma^2}{2p(\rho_2-1)}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.

\item If $d^2 > \frac{2\sigma^2}{p (\rho_2 - 1)}$, then we have the following transition depending on $\rho_1$:
		\begin{itemize}
			\item If $\rho_1 < \frac{(\rho_2-2)\sigma^2}{(1 + {\rho_1}^{-0.5})^4(\rho_2 - 1) pd^2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $\rho_1 > \frac{(\rho_2-2) \sigma^2}{(1 - {\rho_1}^{-0.5})^4 (\rho_2 - 3) pd^2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
\end{proposition}

The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}.
We compare the model shift bias and the amount of reduced variance of $\hat{\beta}_t^{\MTL}$.
An intuitive interpretation of Proposition \ref{prop_data_size} is that:
i) If the two task models are sufficiently similar (as specified under the first bullet), adding the source task always provides positive transfer;
ii) Otherwise, as we increase the number of source task data points, the transfer is positive initially, but transitions to negative eventually.
We leave the proof of Proposition \ref{prop_data_size} to Appendix \ref{app_proof_31}.

\textbf{Covariate shift.}
So far we have considered settings where $\Sigma_1 = \Sigma_2$.
This setting is relevant for multi-class image classification settings, where different tasks share the same input features.
In general, the covariance matrices of the two tasks may be different, e.g. in text classification.
In this part, we use our tools to provide a case study on the effect of applying multi-task learning for two tasks when $\Sigma_1 \neq \Sigma_2$.

For this setting, covariate shift is captured by the matrix $M = \Sigma_1^{1/2} \Sigma_2^{-1/2}$.
We ask: is it better to have $M$ as being close to identity, or should $M$ involve varying levels of singular values?
Understanding this question has implications for applying normalization methods in multi-task learning \cite{LV19,CBLR18,YKGLHF20}.
Our result shows that if $n_1$ is much larger than $n_2$, then the optimal $M$ matrix should be proportional to identity, under certain assumptions on its range of singular values (to be formulated in Proposition \ref{prop_covariate}).
On the other hand, if $n_1$ is comparable or even smaller than $n_2$, we show an example where having ``complementary'' covariance matrices is better performing than having the same covariance matrices.

	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, we assume an upper bound on the scale of $M$.
	Consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	For the task models, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ has i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is large enough compared to $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized approximately within the family of $\cS_{\mu}$ when $M = \mu\id$.

\begin{proposition}\label{prop_covariate}
	In the simplified model, assume that $\rho_1 > 3$ and $\rho_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$. %As a parameter of $M \in \cS_{\mu}$,
The test error $\te(\hat{\beta}_t^{\MTL})$, as a function of $M\in \cS_{\mu}$, is minimized at the identity matrix $\mu\id$ approximately by
$1+ \bigo{\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}}$.
%\be\label{formular_covariate0} (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \left[1+ C\left(\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}\right)\right]\cdot\min_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) ,\ee
%where the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $C_1$, but otherwise does not depend on $c_1$ and $c_2$.
	%is minimized when $M = \mu\id$.
\end{proposition}
Proposition \ref{prop_covariate} shows that when $n_1\gg n_2$, $te(\hat{\beta}_t^{\MTL})$ is minimized when $\Sigma_1$ and $\Sigma_2$ are proportional to each other.
In other words, there is no covariate shift between the source task data and target task data.
This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.

\begin{example}\label{ex_complement}
	Within the setting of Example \ref{ex_cov_family}, we compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	For simplicity, we assume that $d = 0$.
	Hence the two tasks have the same model parameters.

@@

\subsection{Implications on the Benefits of Multi-task Learning}

>>>>

\begin{proposition}\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
		\item[a)] If $pd^2 \le \frac{1}{2}\cdot \frac{\sigma^2}{\rho_2-1}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ for any $\rho_1 > 40$.
		\item[b)] If $pd^2 > 2\cdot\frac{\sigma^2}{\rho_2 - 1}$, then we have the following transition depending on $\rho_1$:
		\begin{itemize}
			\item If $\rho_1 \cdot p < \frac{(\rho_2-2)\sigma^2}{(1 + {\rho_1}^{-0.5})^4(\rho_2 - 1) d^2 - \sigma^2 / p}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $\rho_1 \cdot p > \frac{(\rho_2-2) \sigma^2}{(1 - {\rho_1}^{-0.5})^4 (\rho_2 - 3) d^2 - \sigma^2 / p}$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
\end{proposition}
The intuition behind Proposition \ref{prop_data_size} is as follows.
a) If $\norm{\beta_1 - \beta_2}^2 \approx pd^2$ is sufficiently small, then adding any amount of labeled data from the source task always provides positive transfer;
ii) Otherwise, as source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

\textbf{Covariate shift.}
So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
In general, the covariance matrices of the two tasks may be different such as in text classification.
In this part, we consider what happens when $\Sigma_1 \neq \Sigma_2$.

We measure covariate shift by the matrix $M = \Sigma_1^{1/2} \Sigma_2^{-1/2}$.
We ask: is it better to have $M$ as being close to identity, or should $M$ involve varying levels of singular values?
Understanding this question has implications for applying normalization methods in multi-task learning \cite{LV19,CBLR18,YKGLHF20}.
We show that if $n_1$ is much larger than $n_2$, then the optimal $M$ matrix should be proportional to identity, under certain assumptions on its range of singular values (to be formulated in Proposition \ref{prop_covariate}).
On the other hand, if $n_1$ is comparable or even smaller than $n_2$, we show an example where having ``complementary'' covariance matrices is better performing than having the same covariance matrices.

	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	Similar to the isotropic model, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1$ equals $\beta_2$ plus i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is much larger than $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized at $M = \mu\id$ among all matrices in $\cS_{\mu}$.

\begin{proposition}\label{prop_covariate}
	In the setting described above, assume that $\rho_1 > 3$ and $\rho_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}}} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,

\todo{} To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.
\begin{example}\label{ex_complement}
	In the setting of Proposition \ref{prop_covariate}, we compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	For simplicity, we assume that $d = 0$.
	Hence the two tasks have the same model parameters.

@@

\subsection{Implications on the Benefits of Multi-task Learning}\label{sec_benefit}
